### 2014-11-08 paper/main.tex
added rmd file looking at FCR in Efron's example

Here we consider one setting in which FDR-based methods are widely used:
an experimenter has measured a large numbers of ``effects", each with an associated
measurement precision (standard error), 
and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence).
This setting arises commonly in many areas of statistics, and particularly in genomics applications.
For example, in genomics
a common goal is to identify ``differentially-expressed genes" -- that is, genes
whose mean expression (activity) level differs in two conditions.

@@

Many tools have been developed for this problem \cite{}, and several are widely used.
Here we argue that these existing tools have sub-optimal features that can be improved on, at least in some circumstances.
In particular we argue that existing tools i) fail to take proper account of variations in the measurement precision of different effects;
ii) make, implicitly, unrealistic assumptions about the distribution of the non-zero effects. We provide methods
to address these problems, implemented in an R package \ashr (short for ``Adaptive Shrinkage") and demonstrate that these methods can lead to a substantial reduction in estimated FDRs (or, equivalently,
a substantial increase in the number of discoveries at a given FDR).

Although we motivate our tools by FDR, we are providing generic tools for shrinkage, and
these could be useful in settings such as wavelets.
The potential for shrinkage-based estimation to
address the multiple comparisons problems tackled by the FDR
has been highlighted
previously, including \cite{greenland1991empirical},
\cite{efron2008microarrays} and \cite{gelman2012we}. [Note, check also Louis, JASA, 1984]

@@

Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements $j=1,\dots,J$ to improve inference for each individual $\beta_j$. By improved inference, we mean both
improved average accuracy of point estimates, which is
the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty.

By ``adaptive" shrinkage we 
have two key properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified. Second, the amount of shrinkage undergone by each $\bhat_j$ will depend on the standard error $s_j$: measurements with high standard error will undergo more shrinkage than measurements with low standard error.

Given that shrinkage estimation is widely recognized as a powerful tool, there are surprisingly few software packages for performing the simplest type of shrinkage estimation considered here. (There are more packages for the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
The only package we have found that provides anything similar
to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of \ashr are that it
i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).
NOTE: should emphasize these differences in the examples.

@@

@@

 In outline, we use a hierarchical model to combine information across measurements, with the aim of improving both accuracy and precision of estimates. 
 Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
 \begin{equation} \label{eqn:beta}
 \beta_j \sim g(\cdot),
 \end{equation}
 where $g$ will be estimated from
 the data.  Our methods make the following key assumptions, which distinguish them from most previous work in this area:
   \begin{enumerate}
  \item[A1:] The distribution $g(\cdot)$ is unimodal. (We assume here that the mode is at 0, although 
this could be relaxed.)  
 \item[A2:] The likelihood $L(\beta_j) := p(D_j | \beta_j)$ can be approximated by $p(\bhat_j | \beta_j, \shat_j)$, given by \ref{eqn:betahat-t}.
Equivalently
  \begin{equation} \label{eqn:lik}
 \hat{L}_j(\beta_j) \propto t_\nu((\beta_j-\bhat_j)/\shat_j)
 \end{equation}
 where $t_\nu$ denotes the density of the standard $t$ distribution on $\nu$ degrees of freedom.
%which, in light of (\ref{eqn:betahat}), is equivalent to
 % \begin{equation} \label{eqn:lik2}
% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
 %\end{equation}
 \end{enumerate}

@@

Interestingly, although the unimodal assumption A1 seems natural, much previous related work on estimating False Discovery Rates  
has assumed, either explicitly or implicitly, that the effect sizes are multimodal (see Results for more details). 
In contrast, almost all analogous work in large-scale regression assumes
a unimodal prior distribution for the effect sizes, common choices being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.
One of our aims here is to emphasize potential benefits of also making the unimodal assumption in the FDR context.

Assumption A2 can derived from two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$. The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$. Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$. If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$. See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.

@@

While Assumption A2 has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
 that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
there is an important difference: here, observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
other parameters, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference. 
Because of this the likelihood-based approach can take better account of the informativeness of each measurement, as illustrated in Results below.

@@

Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
\begin{equation}
\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
\end{equation}
Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
were to make our best guess. (Note that we count it as an error to state that $\beta$ is positive or negative when it is truly zero.)
To illustrate, suppose for concreteness
that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
our best guess would be that $\beta$ is negative, and the probability that we have
made an error in sign would be 0.05. We note that the idea of focussing on tail areas, rather than point null hypotheses,
has a long history (e.g. Altham? others?).

@@

>>>>

Here we consider one setting in which FDR-based methods are widely used:
an experimenter has measured a large numbers of ``effects", each with an associated
measurement precision (standard error), 
and the goal is to identify which of the effects we can declare to be non-zero (with some prescribed level of confidence), and also
to provide estimates -- both point estimates and interval estimates -- for these effects.
This setting arises commonly in many areas of statistics, and particularly in genomics applications.
For example, in genomics
a common goal is to identify ``differentially-expressed genes" -- that is, genes
whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
expression level between the two conditions, often referred to as the ``log-fold change". 
%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.

@@

Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and several are widely used.
Here we present new tools that bring two important new ideas to bear on the problem.
First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
a substantial increase in the number of discoveries at a given FDR).

Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests, 
provides a general tool for shrinkage estimation. Shrinkage estimation is a powerful tool, with many potential applications, including for example wavelet denoising, 
where shrinkage of wavelet coefficients is used to smooth signals. 
However, in contrast to FDR methods, shrinkage methods are not so widely used in practice, at least in the kinds
of genomics settings I am familiar with. One possible reason for this is that most work on shrinkage-based estimation has focussed
on point estimates, rather than on interval estimates which are often of more direct practical relevance. Another contributing
factor may be the paucity of general-purpose software packages for this kind of shrinkage estimation.
 (There are, perhaps, more packages for shrinkage esitmation in the more complex setting of covariance estimation, where shrinkage is perhaps still more important.)
Our work here addresses both these issues, providing a general-purpose software tool that uses shrinkage ideas to improve both
point and interval estimates.

By ``adaptive" shrinkage we 
have two properties in mind. First, the appropriate amount of shrinkage is determined from the data, rather than being pre-specified.
Second, the amount of shrinkage undergone by each measurement will depend on its precision, or standard error: measurements with high standard error will undergo more shrinkage than measurements with low standard error.  
Although shrinkage-based estimation can be motivated in various ways, our key goal here is to combine information across the multiple measurements to improve inference for each individual effect. By improved inference, we mean both
improved average accuracy of point estimates, which is
the traditional focus of shrinkage-based methods, \emph{and} improved assessments of uncertainty, such as credible intervals.

The potential for shrinkage-based estimation to
address the multiple comparisons problems tackled by FDR-based methods
has been highlighted
previously, including \cite{greenland1991empirical},
\cite{efron2008microarrays} and \cite{gelman2012we}. [Note, check also Louis, JASA, 1984] 
Furthermore,

@@

Given that shrinkage estimation is widely recognized as a powerful tool, there are

The only package we have found that provides anything similar
to the functionality provided here is \mixfdr \cite{muralidharan2010empirical}. Compared with \mixfdr, the key features of \ashr are that it
i) focuses on allowing for variation in the standard deviation of each observation; ii) constrains the underlying density to be unimodal (and possibly symmetric).

@@

Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?

@@

 In outline, we use an ``Empirical Bayes" hierarchical model to combine information across measurements and perform inference.
 Here performing inference means obtaining an (approximate) posterior distribution for each effect $\beta_j$,
which can be used to estimate False Discovery Rates, and to produce both point and interval estimates for $\beta_j$.
 Although ``multiple comparisons" is usually presented as a ``problem", our emphasis here is that the multiple measurements
 actually provide an opportunity: the hierarchical model combines information across measurements to improve both accuracy and precision of estimates.

 Specifically we assume that the effects $\beta_j$ are independent and identically distributed from some unknown distribution $g(\cdot)$,
 \begin{equation} \label{eqn:beta}
 \beta_j \sim g(\cdot),
 \end{equation}
 where $g$ will be estimated from
 the data.   
  Our methods further make the following key assumptions, which distinguish them from most previous work in this area:
   \begin{enumerate}
  \item[A1:] The distribution $g(\cdot)$ is unimodal. (We assume here that the mode is at 0, although 
this could be relaxed.)  
 \item[A2:] The likelihood $L(\beta) := p(D | \beta)$ can be approximated by $\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
 $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by (\ref{eqn:betahat-t}).
Equivalently
  \begin{equation} \label{eqn:lik}
 \hat{L}(\beta) \propto \prod_j t_\nu((\beta_j-\bhat_j)/\shat_j)
 \end{equation}
 where $t_\nu$ denotes the density of the standard $t$ distribution on $\nu$ degrees of freedom.
%which, in light of (\ref{eqn:betahat}), is equivalent to
 % \begin{equation} \label{eqn:lik2}
% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
 %\end{equation}
 \end{enumerate}

@@

Interestingly, although the unimodal assumption A1 seems natural, much previous related work on estimating False Discovery Rates  
has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}. 
In contrast, almost all analogous work in large-scale regression assumes the effects (regression coefficents) to have a unimodal prior distribution - 
common choices being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors.
One of our aims here is to emphasize the considerable practical benefits of making the unimodal assumption in the FDR context.

Assumption A2 can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$. The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$. Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$. If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$. See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.

@@

While Assumption A2 has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
 that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
there is an important difference: here, observations with a large $\shat_j$ have an essentially flat likelihood (\ref{eqn:lik}), and so do not affect inference for
$g$ or the other effects, whereas when modeling the $Z$ scores directly, observations with large $\shat_j$ produce $Z$ scores with an $N(0,1)$ distribution that {\it do} affect inference.  Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.

@@

Motivated by this, we define the local False Sign Rate (lfsr) for observation $j$ as 
\begin{equation}
\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
\end{equation}
Thus $\lfsr_j$ gives the probability that we would get the sign of $\beta_j$ wrong if we
were to make our best guess. (Note that we count it as an error to state that $\beta$ is positive or negative when it is truly zero.)
To illustrate, suppose for concreteness
that the minimum is achieved by the first term, $p(\beta_j \geq 0| \bhat, s)=0.05$ say. Then
our best guess would be that $\beta$ is negative, and the probability that we have
made an error in sign would be 0.05.

We note that the idea of focussing on tail areas, rather than point null hypotheses,
has a long history (e.g. Altham? others?). For example, 
Tukey, p32, "the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?""
More recently Gelman.

@@

Comparisons with Efron, whose 1-group model is effectively the approach we have taken here.
He says ``deconvolution
calculations are notoriously tricky and I have not
been able to produce a stable general algorithm." Here our claim is that the unimodal constraint is enough to produce that
general stable algorithm.
